Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid

### 2016-02-11 Article.tex
intro modifs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Abstract
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century. It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with the use of small renewable generators and a proper efficient information system. All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation. In this paper, we propose an algorithm that forms coalitions of agents, called prosumers, that both produce and consume. It is designed to be used by aggregators that aim at selling the aggregated surplus of production of the prosumers they control. We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each prosumer. This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors. As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk. The proposed graph based algorithm takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability. We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
\end{abstract}

@@

The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system. The rigid top-down centralized architecture of today’s power grids is expected to evolve in more distributed and flexible systems. Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition. Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages [5]. 
In this paper, we focus on these "nodes" in the distribution network that can produce and consume electricity. More precisely, we consider agents owning both renewable distributed energy resources (DER) and electrical loads. The production of an agent can be used to meet its own demand, but in cases when it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid. Such an agent model is known as a "prosumer" [6]. A key point of this work is to merge the interests of the grid operator with the prosumers in a single utility function. While the prosumers intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the injections.

%The coalitions that we wish to form should thus be both stable and productive. This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation happen.

%More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time. If a contract is bought, the entity commits to injecting, at any time of the contracted period, the announced amount of power, under financial penalties if it fails. Since prosumers use exclusively renewables, contracts come naturally with some amount of risk due to the intermittent nature of most renewables. Using storage as buffers as to reduce risks is a popular approach. Without storage and proper control, the over-producing state of the agents, and thus the power they inject in the grid, might be rather unstable. This is unacceptable for the grid operator that cannot ensure system stability if it has to deal with numerous small unpredictable entities.

We rely on a market representation of the interactions. Coalitions of prosumers estimate a production capacity for an upcoming period and announce it on the market. If some other entity buys the contract, the coalition has to inject exactly the contracted amount of power at any time. Obviously, deviations will cause financial penalties to the coalitions, such that respecting the contracts is paramount. Since prosumers use exclusively renewables, we are particularly concerned with the probability that the available capacity of a coalition falls below its contract value. In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.

We focus on coalitions because it appears as a solution to both the number of prosumers communicating with the operator as well as their high variabilities. Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side \cite{Negeri2012}. On the other hand, it is known that diversification of the assets is a way of minimizing risks when building a portfolio. We thus expect more stable and predictable energy productions for coalitions than for single agents. Nevertheless, all coalitions are not equally stable or productive. Special attention should thus be paid to the aggregation step. Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents. This is a central topic of this paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?

We will see that the variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions. Understanding the correlation relationships among the agents can thus give an indication about what coalitions to form and how much they should sell. More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions. We build a framework in which the system operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to accept. We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}). Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of acceptable risk. In other words, it tries to maximize coalitions profits without considering individual retributions to single agents. Furthermore, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.

@@

Assuming a large penetration of DER, it is expected that the managment of the system would have to evolve in order to copy with the large amount of unpredictable small generators. In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high use of renewables such that power is scheduled in a distributed fashion. Allowing the formation of coalitions whose participants commit to achieving some objective against a reward appears also as a promising approach. Nevertheless, for these structures to be stable toward defection, agents should have an incentive to join, and cooperate. Coalition formation is therefore usually studied through the framework of game theory \cite{Saad2011}. In this direction, \cite{Nguyen2013} introduces an agent-based strategy for managing DER aiming at optimal performance in the electricity supply system.

@@

The contract a given coalition is willing to take depends on its capacity to compensate for under-producing (using batteries, backup generators...), and its risk acceptance. Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works. In order to keep the present paper in a reasonable length, we simplify the contract value selection problem by giving some responsibilities to a third party named the grid operator. The role of the grid operator is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values. More formally, let $ \phi \in [0,1] $ be the reliability threshold fixed by the grid operator as a maximum value for the probability of under-producing. The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $. In the Gaussian example, it implies that coalition $ \{i,j\}$ is announcing :

@@

This is the best contract value that the coalition S can afford for a given stability policy $\phi$ of the grid operator. Figure \ref{fig:Gaussian} shows how $ P_{S}^{CRCT \star} $ evolves according to the reliability parameter $ \phi $. For illustration, the range of $ \phi $ values is shown from 0 to 1, but in practice, only small values of $ \phi $ really make sense : $ \phi = 1 $ for instance means that coalitions can announce absolutely anything since the probability of producing less than any contract value is necessarily less than one by trivial definition of a probability. As visible on figure \ref{fig:Gaussian}, coalitions with high expected productions but presenting a high unpredictability are penalized and can only afford small contracts. The grid operator also specifies a lower bound $ P^{MIN} $ on the contract values as not to overload the market with unrealistic small coalitions. We thus characterized a valid coalition as one satisfying the two conditions :

@@

\begin{itemize}
	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
\end{itemize}

@@

A key point for the coalitions, besides stability and productivity, is their resilience. The resilience of a system refers to its ability to perform its tasks when subject to failures of its components. We consider here the case of random failures of the power electronics of some agents that has the consequence of preventing them from participating in the market. Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when some of its internal agents are removed. According to our model, the grid operator specified two thresholds ($P^{MIN}$ and $ \phi $) such that the power injected by every coalition is constrained : $ P_{S}^{CRCT} \in [P^{MIN}, P_{S}^{CRCT \star}] $. As long as a coalition can propose a contract value higher than $ P^{MIN} $, it is valid and allowed to enter the energy market. We define the resilience of a coalition S as the probability that S produces more than the $ P^{MIN} $ threshold : $ \mathcal{R}_{S} = Pr[P_{S} >= P^{MIN}] = 1 - Pr[P_{S} < P^{MIN}] $. And we extend this measure to the coalition structures :

@@

In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production to the grid operator. By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions. The power that a coalition is able to propose on the market is therefore related to production and stability. The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents. We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation. A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results. We compared these results with random samplings, and an opposite strategy that clusters correlated agents together. We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility. Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions. Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations. We therefore studied how the coalitions remain on the market when their agents fail randomly. We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents. A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators as to propose a prosumer aggregation model closer to real conditions.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Abstract
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century. It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with the use of small renewable generators and a proper efficient information system. All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation. In this paper, we propose an algorithm that forms coalitions of agents, called prosumers, that both produce and consume. It is designed to be used by aggregators that aim at selling the aggregated surplus of production of the prosumers they control. We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each prosumer. This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors. As aggregators are bound to the market operator by a contract, they seek to maximize their offer while minimizing their risk. The proposed graph based algorithm takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability. We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
\end{abstract}

@@

The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system. The rigid top-down centralized architecture of today’s power grids is expected to evolve in more distributed and flexible systems. Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition. Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages.

In this paper, we consider agents in the distribution network that own both renewable distributed energy resources (DER) and loads. We will refer to these agents as prosumers \cite{Rathnayaka2012} in the future. If a prosumer produces more than he consumes, he might be willing to sell power. We thus assume the existence of an energy market where producers post a production capacity for an upcoming period, and buyers decide what bids to purchase. When the production is stochastic, as it can be with DER, this approach gets more complicated and risk should be quantified and taken into account. Rather than having a risk dependant price, we propose here a safer and simpler model. We assume the existence of a market operator who constrains the market entrance. A key point of this work is to merge the interests of the market operator with the prosumers in a single utility function. While the prosumers intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply.

If a buyer purchases a contract, the corresponding producer has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates. Since prosumers use renewables, we are concerned with the probability that the production falls below the contract value. In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low. Single prosumers might be authorized to enter such market, but it is unlikely that a single agent can propose enough power with enough reliability.

Therefore, we use a coalitional framework in order to model aggregators of prosumers. It is known that diversification of the assets is a way of minimizing risks when building a portfolio. We thus expect more stable and predictable energy productions for aggreagtors than for single agents. Nevertheless, all coalitions are not equally stable or productive. Attention should thus be paid to the aggregation step. Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents. This is a central topic of this paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?

We will see that the variability in the production of the aggregators can be quantified to a certain extent by the correlation among the agents within the coalitions. Understanding the correlation relationships among the agents can thus give an indication about what coalitions to form and how much they should sell on the market. More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions. We build a framework in which the market operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to accept. We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}). Rather than maximizing the profit of each agent in the grid (prosumers and operator), our algorithm tries to form the most productive coalitions given a maximum amount of acceptable risk. In other words, it tries to maximize aggregators' profits without considering individual retributions to single agents. Furthermore, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.

@@

Assuming a large penetration of DER, it is expected that the managment of the system would have to evolve as to copy with the large amount of unpredictable small generators. In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high use of renewables such that power is scheduled in a distributed fashion. Allowing the formation of coalitions whose participants commit to achieving some objective against a reward appears also as a promising approach. Nevertheless, for these structures to be stable toward defection, agents should have an incentive to join, and cooperate. Coalition formation is therefore usually studied through the framework of game theory \cite{Saad2011}. In this direction, \cite{Nguyen2013} introduces an agent-based strategy for managing DER aiming at optimal performance in the electricity supply system.

@@

The contract a given coalition is willing to take depends on its capacity to compensate for under-producing (using batteries, backup generators...), and its risk acceptance. Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works. In order to keep the present paper in a reasonable length, we simplify the contract value selection problem by giving some responsibilities to a third party named the market operator. His role is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values. More formally, let $ \phi \in [0,1] $ be the reliability threshold fixed by the market operator as a maximum value for the probability of under-producing. The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $. In the Gaussian example, it implies that coalition $ \{i,j\}$ is announcing :

@@

This is the best contract value that the coalition S can afford for a given stability policy $\phi$ of the market operator. Figure \ref{fig:Gaussian} shows how $ P_{S}^{CRCT \star} $ evolves according to the reliability parameter $ \phi $. For illustration, the range of $ \phi $ values is shown from 0 to 1, but in practice, only small values of $ \phi $ really make sense : $ \phi = 1 $ for instance means that coalitions can announce absolutely anything since the probability of producing less than any contract value is necessarily less than one by trivial definition of a probability. As visible on figure \ref{fig:Gaussian}, coalitions with high expected productions but presenting a high unpredictability are penalized and can only afford small contracts. The market operator also specifies a lower bound $ P^{MIN} $ on the contract values as not to overload the market with unrealistic small coalitions. We thus characterized a valid coalition as one satisfying the two conditions :

@@

\begin{itemize}
	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
	\item \textbf{The market operator :} market entrance policy $ (P^{MIN},\phi) $,
	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
\end{itemize}

@@

A key point for the coalitions, besides stability and productivity, is their resilience. The resilience of a system refers to its ability to perform its tasks when subject to failures of its components. We consider here the case of random failures of the power electronics of some agents that has the consequence of preventing them from participating in the market. Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when some of its internal agents are removed. According to our model, the market operator specified two thresholds ($P^{MIN}$ and $ \phi $) such that the power injected by every coalition is constrained : $ P_{S}^{CRCT} \in [P^{MIN}, P_{S}^{CRCT \star}] $. As long as a coalition can propose a contract value higher than $ P^{MIN} $, it is valid and allowed to enter the energy market. We define the resilience of a coalition S as the probability that S produces more than the $ P^{MIN} $ threshold : $ \mathcal{R}_{S} = Pr[P_{S} >= P^{MIN}] = 1 - Pr[P_{S} < P^{MIN}] $. And we extend this measure to the coalition structures :

@@

In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production on the energy market. By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions. The power that a coalition is able to propose on the market is therefore related to production and stability. The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents. We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation. A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results. We compared these results with random samplings, and an opposite strategy that clusters correlated agents together. We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility. Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions. Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations. We therefore studied how the coalitions remain on the market when their agents fail randomly. We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents. A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators as to propose a prosumer aggregation model closer to real conditions.
